---
title: "p8105_hw6_tj2383"
author: "Tanvi Jain"
date: "11/16/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mgcv)
library(modelr)
set.seed(1)
```

###Problem 1

Importing the homicide dataset from the Washington Post.
```{r}
homicide_data = read.csv("./data/homicide-data.csv")
```


In this code chunk I create a `city_state` variable by combining the `city` and `state` variables and a binary variable indicating whether the homicide is solved. I omitted the cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. I also omitted Tulsa, AL – this is a data entry mistake. Next I modifiy `victim_race` to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.
```{r}
homicide_data_clean = 
  homicide_data %>% 
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  select(-city, -state) %>% 
  mutate(case_resolution = as.numeric(disposition == "Closed by arrest")) %>% 
  filter(city_state != "Dallas, TX" & city_state != "Phoenix, AZ" & city_state != "Kansas City, MO" & city_state != "Tulsa, AL") %>% 
  
  mutate(binary_race = case_when(
    victim_race == "Asian" ~ "non-white",
    victim_race == "Black" ~ "non-white",
    victim_race == "Hispanic" ~ "non-white",
    victim_race == "Other" ~ "non-white",
    victim_race == "Unknown" ~ "non-white",
    victim_race == "White" ~ "white"),
    
    victim_race = factor(binary_race, levels = c("white", "non-white")),
    victim_age = as.numeric(as.character(victim_age))
  )
```


In this code chunk I use the glm function to fit a logistic regression with `case_resolution` as the outcome and `victim age`, `sex`, and `race` as predictors for Baltimore, MD. I then saved the output of glm as an R object; applied the broom::tidy to this object; and obtained the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing black victims to white victims, keeping all other variables fixed.
```{r}
fit_logistic_baltimore = 
  homicide_data_clean %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(case_resolution ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 

fit_logistic_baltimore %>% 
  broom::tidy() %>% 
  janitor::clean_names() %>% 
  mutate(OR = exp(estimate),
         lower_95_limit = exp(estimate - (1.96 * std_error)),
         upper_95_limit = exp(estimate + (1.96 * std_error))) %>%
  filter(term == "victim_racenon-white")
```
The odds of having a solved case among those who identify as non-white is 0.44 times the odds of having a solved case among those who identify as white and we are 95% confident that these odds are between 0.31 and 0.62 adjusting for sex and age.


In this code chunk I ran glm for each of the cities in the dataset, and extracted the adjusted odds ratio and 95% CI for solving homicides comparing black victims to white victims. I did this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest.
```{r}
nested_glm =
  homicide_data_clean %>% 
  group_by(city_state) %>% 
  nest() %>% 
  mutate(regression_city_state = map(data, ~lm(case_resolution ~ victim_age + victim_sex + victim_race, data = .x)),
         regression_city_state = map(regression_city_state, broom::tidy)) %>% 
  select(-data)

homicides = nested_glm %>% 
  unnest() %>% 
  janitor::clean_names() %>% 
  mutate(OR = exp(estimate),
         lower_95_limit = exp(estimate - (1.96 * std_error)),
         upper_95_limit = exp(estimate + (1.96 * std_error))
  ) %>% 
  filter(term == "victim_racenon-white") %>%
  select(city_state, OR, lower_95_limit, upper_95_limit)
```

This is a dataframe consisting of the estimated ORs and CIs for each city.
`r homicides %>% knitr::kable()`


In this code chunk I created a plot that shows the estimated ORs and CIs for each city. I organized cities according to estimated OR.
```{r}
homicides %>% 
  mutate(city_state = forcats::fct_inorder(city_state)) %>% 
  ggplot(aes(x = city_state, y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_95_limit, ymax = upper_95_limit)) +
  labs(
    x = "city, state",
    y = "ORs",
    title = "ORs and 95% CIs for solved cases comparing non-whites to whites"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) 
```

According to this scatter plot, majority of the ORs appear to be below 1 with the exception of Birmingham, AL, Durham, NC, and Tampa, FL, indicating that majority of the cities have decreased odds of a case being solved for non-white people compared to white people, except for the three cities with a greater than 1 odds ratio. However, the null value of 1 falls in the 95% CIs for these three cities indicating that their ORs aren't significant. Notably, there are also some wider confidence intervals, indicating less accuracy regarding the estimate's true value.


###Problem 2

In this code chunk I imported the child birthweight data and and cleaned the data for regression analysis (i.e. specified reference levels for categorical variables, converted numeric to factor where appropriate, and checked for missing data).
```{r}
birthweight_data = read.csv("./data/birthweight.csv") %>% 
  janitor::clean_names()

missing_values = sum(is.na(birthweight_data))

birthweight_data = 
  birthweight_data %>% 
  mutate(
    babysex = case_when(
      babysex == 1 ~ "male",
      babysex == 2 ~ "female"),
    babysex = fct_relevel(babysex, "male"),
    frace = case_when(
      frace == 1 ~ "white",
      frace == 2 ~ "black",
      frace == 3 ~ "asian",
      frace == 4 ~ "puerto_rican",
      frace == 8 ~ "other",
      frace == 9 ~ "unk"),
    frace == fct_relevel(frace, "white"),
    malform = case_when(
      malform == 0 ~ "absent",
      malform == 1 ~ "present"),
    malform == fct_relevel(malform, "absent"),
    mrace = case_when(
      mrace == 1 ~ "white",
      mrace == 2 ~ "black",
      mrace == 3 ~ "asian",
      mrace == 4 ~ "puerto_rican",
      mrace == 8 ~ "other"),
    mrace == fct_relevel(mrace, "white")
  )
```
This dataset contains no missing values.


In this code chunk I created a linear regression model for birthweight.
```{r}
fit_linear_birthweight = 
  birthweight_data %>% 
  lm(bwt ~ momage + babysex + fincome + gaweeks + malform + mrace, data = .)
```
This model is based on my hypothesis for key maternal sociodemographic and biologic indicators that influence infant birthweight. This model is testing the association between mother's age at delivery on baby's birthweight adjusting for family income, baby's sex, gestational age, presence of malformations, and mother's race.


The following is a plot of model residuals against fitted values. 
```{r}
modelr::add_residuals(birthweight_data, fit_linear_birthweight) 
modelr::add_predictions(birthweight_data, fit_linear_birthweight)

birthweight_data %>% 
  modelr::add_residuals(fit_linear_birthweight) %>% 
  modelr::add_predictions(fit_linear_birthweight) %>% 
  ggplot(aes(x = pred, y = resid)) + geom_point() +
  theme_bw()
```


In this code chunk I compared my model to two other models.
```{r}
model_1 =
  birthweight_data %>% 
  lm(bwt ~ blength + gaweeks, data = .)

model_2 = 
  birthweight_data %>% 
  lm(bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength+babysex + bhead*blength*babysex, data = .)

cv_birthweight = 
  crossv_mc(birthweight_data, 100)

cv_birthweight = 
  cv_birthweight %>% 
  mutate(fit_linear_birthweight = map(train, ~lm(bwt ~ momage + babysex + fincome + gaweeks + malform + mrace, data = .x)),
         model_1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         model_2 = map(train, ~lm(bwt ~ bhead + blength + babysex + bhead*blength*babysex + bhead*blength + bhead*babysex + blength+babysex, data = .x))) %>% 
  mutate(rmse_fit_linear_birthweight = map2_dbl(fit_linear_birthweight, test, ~rmse(model = .x, data = .y)),
         rmse_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
         rmse_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y)))

```


