---
title: "p8105_hw6_tj2383"
author: "Tanvi Jain"
date: "11/16/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
set.seed(1)
```

###Problem 1

Importing the homicide dataset from the Washington Post.
```{r}
homicide_data = read.csv("./data/homicide-data.csv")
```


In this code chunk I create a `city_state` variable by combining the `city` and `state` variables and a binary variable indicating whether the homicide is solved. I omitted the cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. I also omitted Tulsa, AL – this is a data entry mistake. Next I modifiy `victim_race` to have categories white and non-white, with white as the reference category. Be sure that victim_age is numeric.
```{r}
homicide_data_clean = 
  homicide_data %>% 
  mutate(city_state = paste(city, state, sep = ", ")) %>% 
  select(-city, -state) %>% 
  mutate(case_resolution = disposition == "Closed by arrest") %>% 
  filter(city_state != "Dallas, TX" & city_state != "Phoenix, AZ" & city_state != "Kansas City, MO" & city_state != "Tulsa, AL") %>% 
  
  mutate(binary_race = case_when(
    victim_race == "Asian" ~ "non-white",
    victim_race == "Black" ~ "non-white",
    victim_race == "Hispanic" ~ "non-white",
    victim_race == "Other" ~ "non-white",
    victim_race == "Unknown" ~ "non-white",
    victim_race == "White" ~ "white"),
    
    victim_race = factor(binary_race, levels = c("White", "Non-White")),
    victim_age = as.numeric(as.character(victim_age))
  )
```


